---
title: "lesson3_problems"
output: html_document
---

## Load diamonds data set
```{r}
setwd("/Users/brianchase/R")
library(ggplot2)
data(diamonds)

summary(diamonds)

?diamonds

```

## Create a histogram of the price of
## all the diamonds in the diamond data set.
```{r}
qplot(x = price, data = diamonds)
```

## Describe shape and center. include summary statistics like mean and median.
```{r}
summary(diamonds$price)
```

## DIamond counts
How many diamonds cost less than $500
```{r}
sum(diamonds$price < 500)
```

How many diamonds cost less than $250
```{r}
sum(diamonds$price < 250)
```

How many diamonds cost more than $15000
```{r}
sum(diamonds$price >= 15000)
```


## Exploring the largest peak in the price histogram
```{r}
qplot(x = price, data = diamonds, binwidth=1) +
  scale_x_continuous(limits = c(500, 1000), breaks =seq(500,1000,50))

sum(diamonds$price == 605)
```

## Price by Cut histograms (use multiple panels)
```{r}
qplot(x = price, data = diamonds) +
  facet_wrap(~cut)
```

## Finding lowest, highest, lowest median price based on cut
```{r}
by(diamonds$price, diamonds$cut, min)

by(diamonds$price, diamonds$cut, max)

by(diamonds$price, diamonds$cut, median)
```

## histograms with different ranges in y axis (This shows that fair and good are similar distributions)
```{r}
qplot(x = price, data = diamonds) +
  facet_wrap(~cut, scales="free_y")
```

## Price per carat cut
```{r}
qplot(x = price/carat, data = diamonds, bins=500) +
  facet_wrap(~cut, scales="free_y") + 
  scale_x_log10()
```

## Investigate by color with box plot and summary
```{r}
qplot(x = color, y = price, 
      data = diamonds,
      geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 8000))

by(diamonds$price, diamonds$color, summary)

```

What is the price range for the moiddle 50% of diamonds (color D)?
first quartile? 911
third quartile? 4214
Same foar color J:
first quartile? 1860
third quartile? 7695

What is IQR? Q3 - Q1

c) IQR D color: 4214-911 = 3,303
d) IQR J color: 7695-1860 = 5,835

## Price per carat across different colors box plot
```{r}
qplot(x = color, y = price/carat, 
      data = diamonds,
      geom = 'boxplot') +
  coord_cartesian(ylim = c(2000, 6000))

by(diamonds$price / diamonds$carat, diamonds$color, summary)
```

## Carat Frequenct Polygon : What carat size has a count greater than 2000?
```{r}
library(gridExtra)

p1 <- qplot(x = carat, data = diamonds, binwidth = 0.01,
      geom = 'freqpoly') + 
  scale_x_continuous(limits = c(0.3, 0.35), breaks =seq(0.3,0.35,0.01))

p2 <- qplot(x = carat, data = diamonds, binwidth = 0.01,
      geom = 'freqpoly') + 
  scale_x_continuous(limits = c(1, 1.05), breaks =seq(0,1.05,0.01))

grid.arrange(p1,p2, ncol=1)
```

tidyr -a package that reshapes the layout of your data

dplyr - a package that helps you transform tidy, tabular data

[Data wrangling in R presentation](https://s3.amazonaws.com/udacity-hosted-downloads/ud651/DataWranglingWithR.pdf)
[Data wrangling cheat sheet](http://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf)
[other r studio cheat sheets](http://www.rstudio.com/resources/cheatsheets/)

## Load gapminder data on Median Age
```{r}
age_data <- read.csv('data_median_age.csv')

summary(age_data$X1950)

qplot(x = X1950,  
      data = age_data, binwidth = 1) +
  scale_x_continuous(limits = c(15, 34), breaks =seq(15,34,1))

summary(age_data$X2015)

qplot(x = X2015,  
      data = age_data, binwidth = 1) +
  scale_x_continuous(limits = c(14, 47), breaks =seq(14,47,1))

```

Loaded data with median ages based on countries and date (dates also include future projections)

Looking at this data, I notice that there are now three group in the histogram (17-23), (25-31), (36-43)

```{r}
age_data_2015_17to23 <- subset(age_data, age_data$`2015` < 24)
age_data_2015_17to23$`Median age`

age_data_2015_25to31 <- subset(age_data, age_data$`2015` > 25 & age_data$`2015` < 32)
age_data_2015_25to31$`Median age`

age_data_2015_36to43 <- subset(age_data, age_data$`2015` > 35 & age_data$`2015` < 44)
age_data_2015_36to43$`Median age`
```

Plotting the future projection historgram
```{r}
summary(age_data$X2050)
library(gridExtra)


p3 <- qplot(x = `X2050`,  
      data = age_data, binwidth = 1) +
  scale_x_continuous(limits = c(19, 57), breaks =seq(19,57,1))

p2 <- qplot(x = `X1950`,  
      data = age_data, binwidth = 1) +
  scale_x_continuous(limits = c(19, 57), breaks =seq(19,57,1))

p1 <- qplot(x = `X2015`,  
      data = age_data, binwidth = 1) +
  scale_x_continuous(limits = c(19, 57), breaks =seq(19,57,1))

p4 <- grid.arrange(p2, p1, p3, ncol=1)
ggsave('median_age_histograms2.png', plot = p4)
```

Finding another chart to make

qplot(x =, y = ..count../sum(..count..),
      data = subset(pf, !is.na(gender)), binwidth = 10,
      geom = 'freqpoly', color = gender) + 
  scale_x_continuous(limits = c(0, 1000), breaks =seq(0,1000,50)) 


How do you create a plot with the years on the x axis and the median age on the y axis colored by country?
```{r}
#install.packages('tidyr', dependencies=T)
setwd("/Users/brianchase/R")
age_data <- read.csv('data_median_age.csv',header = T,check.names = F)
library(tidyr)
age_data[23] <- NULL
age_data_tidyr <- gather(age_data, "year", "n", 2:22)

qplot(x=age_data_tidyr$year, y=age_data_tidyr$n, data=age_data_tidyr,
      geom = c('point', 'path'), color=age_data_tidyr$`Median age`) +
  guides(color=FALSE)

qplot(x=age_data_tidyr$year, y=age_data_tidyr$n, data=age_data_tidyr,
      geom = c('point', 'path'))
```

Problem is that there are just too many countries in the data set (here I tried to get geom_line to work... turns out you need a group)
```{r}

subset_world <- subset(age_data_tidyr, `Median age` == "AFRICA" | `Median age` == "LATIN AMERICA AND THE CARIBBEAN" | `Median age` == "ASIA" | `Median age` == "EUROPE" | `Median age` == "NORTHERN AMERICA" | `Median age` == "OCEANIA" | `Median age` == "WORLD" )

qplot(x=year, y=n,
      data=subset_world,
      color = `Median age`,
      group = `Median age`,
      geom = c('point', 'line')) +
  scale_y_continuous(limits = c(17, 47), breaks =seq(17,47,1))

ggsave('median_age_chart.png')

summary(subset_world)
```

```{r}
subset_world2 <- subset(age_data_tidyr, `Median age` == "Eastern Africa" | `Median age` == "Western Africa" | `Median age` == "Middle Africa" | `Median age` == "Northern Africa" | `Median age` == "Southern Africa" | `Median age` == "Sub- Saharan Africa" | `Median age` == "AFRICA" | `Median age` == "WORLD" )

qplot(x=year, y=n,
      data=subset_world2,
      color = `Median age`,
      group = `Median age`,
      geom = c('point', 'line')) +
  scale_y_continuous(limits = c(15, 39), breaks =seq(15,39,1))

ggsave('median_age_chart_africa.png')
summary(subset_world2)



```

