---
title: "lesson3"
output: html_document
---
## Load data and initial plot
```{r}
setwd("/Users/brianchase/R")
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep='\t')
qplot(x = friend_count, data = pf)
```

## limit the Axes
```{r}
qplot(x = friend_count, data = pf, xlim = c(0,1000))

qplot(x = friend_count, data = pf) + 
  scale_x_continuous(limits = c(0, 1000))
```

## Adjusting the bin width
```{r}
qplot(x = friend_count, data = pf, binwidth = 25) + 
  scale_x_continuous(limits = c(0, 1000), breaks =seq(0,1000,50)) +
  facet_wrap(~gender)
```

## Ommiting NA gender observations
```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 25) + 
  scale_x_continuous(limits = c(0, 1000), breaks =seq(0,1000,50)) +
  facet_wrap(~gender)
```

## Count number of males and females
```{r}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

## Tenure (Add color to plot)
```{r}
qplot(x = tenure, data = pf, binwidth = 30, 
      color = I('black'), fill = I('#099DD9'))

qplot(x = tenure/365, data = pf, binwidth = .25, 
      color = I('black'), fill = I('#F79420')) + 
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7))
```

## Labling Plots
```{r}
qplot(x = tenure/365, data = pf,
      xlab = 'Number of years using facbeook',
      ylab = 'Number of users in sample',
      color = I('black'), fill = I('#F79420')) + 
  scale_x_continuous(breaks = seq(1, 7, 1), limits = c(0, 7))
```

## User Ages
```{r}
qplot(x = age, data = pf, binwidth = 1,
      xlab = 'Age of users in years',
      ylab = 'Number of users in sample',
      color = I('black'), fill = I('#F79420')) + 
  scale_x_continuous(breaks = seq(0,113,5))
```

## How do you find out of your x variable is continuous or discrete?

## How to find out the max age using summary function
```{r}
summary(pf$age)
```

## Transforming Data
```{r}
qplot(x = age, data = pf)

summary(pf$friend_count)

summary(log10(pf$friend_count + 1))

summary(sqrt(pf$friend_count))
```

[Create Multiple Plots in One Image Output](http://lightonphiri.org/blog/ggplot2-multiple-plots-in-one-graph-using-gridextra)
[Add Log or Sqrt Scales to an Axis](http://docs.ggplot2.org/current/scale_continuous.html)
[Assumptions of Linear Regression](http://en.wikipedia.org/wiki/Linear_regression#Assumptions)
[Normal Distribution](http://en.wikipedia.org/wiki/Normal_distribution)

## Student question: Use scales and multiple plots on one page and install and load gridExtra
1. friend count
2. friend count transformed using log10
3. friend count transformed using sqrt
```{r}
# install.packages('gridExtra') 
library(gridExtra)

p1 <- qplot(x = friend_count, data = pf)
p2 <- qplot(x = log10(friend_count+1), data = pf)
p3 <- qplot(x = sqrt(friend_count), data = pf) 
grid.arrange(p1, p2, p3, ncol=1)
```

[Log Transformations of Data](http://www.r-statistics.com/2013/05/log-transformations-for-skewed-and-wide-distributions-from-practical-data-science-with-r/)

## Alternate solution for transforming data
```{r}
p1 <- ggplot(aes(x = friend_count), data = pf) + geom_histogram()
p2 <- p1 + scale_x_log10()
p3 <- p1 + scale_x_sqrt()

grid.arrange(p1, p2, p3, ncol = 1)
```

## Frequency polygons 
```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), binwidth = 10) + 
  scale_x_continuous(limits = c(0, 1000), breaks =seq(0,1000,50)) +
  facet_wrap(~gender)

qplot(x = friend_count, y = ..count../sum(..count..),
      data = subset(pf, !is.na(gender)), binwidth = 10,
      geom = 'freqpoly', color = gender) + 
  scale_x_continuous(limits = c(0, 1000), breaks =seq(0,1000,50)) 
```

## Equiv ggplot
```{r}
ggplot(aes(x = friend_count, y = ..count../sum(..count..)), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender), binwidth=10) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') + 
  ylab('Percentage of users with that friend count')
```

## Testing ..density..
```{r}
ggplot(aes(x = friend_count, y = ..density.. ) , data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender), binwidth=10) + 
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') + 
  ylab('percentage of users in their own gender group with that friend count')
```

## Student problem: Use a freq polygon to det which gender has more ww_likes
```{r}
qplot(x = www_likes, data = subset(pf, !is.na(gender)), 
      binwidth = 1, geom = 'freqpoly', color = gender) +
  scale_x_continuous(limits = c(0, 200), breaks =seq(0,200,10)) +
  scale_y_log10(limits = c(1,1000), breaks = seq(0,1000,1))
summary(pf$www_likes)
```

# Provided answer
```{r}
qplot(x = www_likes, data = subset(pf, !is.na(gender)),
      geom = 'freqpoly', color = gender) +
  scale_x_continuous() +
  scale_x_log10()
```

## Getting the sum of all likes by gender
```{r}
by(pf$www_likes, pf$gender, sum)

```

## Box Plot
```{r}
qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot', ylim = c(0, 1000))

qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot') +
  scale_y_continuous(limits = c(0, 1000))

qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 1000))
```
Last one is better since it is not dropping data points


[how to useboxplot](http://flowingdata.com/2008/02/15/how-to-read-and-use-a-box-and-whisker-plot/)
[interquartile range or IQR](http://en.wikipedia.org/wiki/Interquartile_range)
[Vis of IQR](http://en.wikipedia.org/wiki/File:Boxplot_vs_PDF.svg)
[intro to descriptive stats exercise: Match box plots](https://www.udacity.com/course/viewer#!/c-ud827/l-1471748603/e-83417918/m-83664035)

[how to interpret box plot](http://flowingdata.com/2008/02/15/how-to-read-and-use-a-box-and-whisker-plot/)

## Zooming in on the box plot (quartile inspection)
```{r}
qplot(x = gender, y = friend_count, 
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 250))

by(pf$friend_count, pf$gender, summary)
```
Median is black line in the center, the 1st and 3 quartile are indicated by the box top and bottom.

## Friendships intiated
```{r}
qplot(x = gender, y = friendships_initiated, 
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot') +
  coord_cartesian(ylim = c(0, 150))

by(pf$friendships_initiated, pf$gender, summary)
```

## Figuring out what number of users checked in with mobile
```{r}
summary(pf$mobile_likes)

summary(pf$mobile_likes > 0)

pf$mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)
```

## What Portion used mobile?
```{r}
sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
```


