---
title: "lesson4_problems"
output: html_document
---

## Quiz: Price vs. x

```{r}
setwd("/Users/brianchase/R")
library(ggplot2)
?diamonds
data("diamonds")
ggplot(aes(x = x, y = price), data = diamonds) + geom_point()
```

## Correlations

```{r}
with(diamonds, cor.test(x, price, method = 'pearson'))

with(diamonds, cor.test(y, price, method = 'pearson'))

with(diamonds, cor.test(z, price, method = 'pearson'))
```

## price vs depth scatterplot

```{r}
ggplot(aes(x = depth, y = price), data = diamonds) + geom_point()
```


## Adjustments - price vs depth

```{r}
summary(diamonds$depth)

ggplot(data = diamonds, aes(x = depth, y = price)) + 
  geom_point(alpha = 1/100) +
  scale_x_continuous(breaks = seq(43, 79, 2))
```

## Correlation - price and depth

```{r}
with(diamonds, cor.test(depth, price, method = 'pearson'))
```


## Create a scatterplot of price vs carat and omit the top 1% of price and carat values.
```{r}
ggplot(aes(x = carat, y = price), data = diamonds) + 
  geom_point() +
  xlim(0, quantile(diamonds$carat, 0.99)) +
  ylim(0, quantile(diamonds$price, 0.99))
```

## Price vs volume
```{r}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z

ggplot(aes(x = volume, y = price), data = diamonds) + 
  geom_point()

```

COunting with plyr package (unload aftewards since plyr conflicts with dplyr)
```{r}
library(plyr)
count(diamonds$volume == 0)
detach("package:plyr", unload=TRUE)
```

## Correlation of price and valume exclude diamonds that have volume 0 or greater than or equal to 800
```{r}
with(subset(diamonds, volume != 0 & volume <=800), cor.test(volume, price))
```

## Adjustments - price vs volume
```{r}
ggplot(aes(x = volume, y = price), data = subset(diamonds, volume != 0 & volume <=800)) + 
  geom_point(alpha = 1/30) +
  geom_smooth(method = 'lm', color = 'red')
```

## Mean price by clarity
```{r}
library(dplyr)

clarity_groups <- group_by(diamonds, clarity)
diamondsByClarity <- summarise(clarity_groups,
                          mean_price = mean(price),
                          median_price = median(price),
                          min_price = min(price),
                          max_price = max(price),
                          n = n())
diamondsByClarity <-arrange(diamondsByClarity , clarity)

head(diamondsByClarity)
```

## Bar Charts of Mean Price
```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

summary(diamonds_mp_by_color$mean_price)

p1 <- ggplot(diamonds_mp_by_color, 
       aes(x = diamonds_mp_by_color$color, y = diamonds_mp_by_color$mean_price)) + 
  geom_bar(stat = "identity") + 
  scale_y_continuous(breaks = seq(0, 5500, 500), limits = c(0, 5500))

p2 <- ggplot(diamonds_mp_by_clarity, 
       aes(x = diamonds_mp_by_clarity$clarity, y = diamonds_mp_by_clarity$mean_price)) + 
  geom_bar(stat = "identity") 

library(gridExtra)
grid.arrange(p1,p2, ncol = 1)
```

## Gapminder example
```{r}
age_data <- read.csv('data_median_age.csv',header = T,check.names = F)

library(tidyr)
library(gridExtra)
age_data[23] <- NULL
age_data_tidyr <- gather(age_data, "year", "n", 2:22)

ggplot(aes(x = as.numeric(year), y = n), data = age_data_tidyr) + 
  geom_point(alpha = 1/5, 
             position = position_jitter(h = 0)) +
  geom_smooth(method = 'lm', color = 'red')

with(age_data_tidyr, cor.test(as.numeric(year), n))

p1 <- ggplot(aes(x = as.numeric(year), y = n), data = age_data_tidyr) + 
  geom_point(alpha = 1/5, 
             position = position_jitter(h = 0)) +
  geom_smooth(method = 'lm', color = 'red')

age_data_2015_17to23 <- subset(age_data, age_data$`2015` < 24)

age_data_tidyr_low <- gather(age_data_2015_17to23 , "year", "n", 2:22)
p2 <- ggplot(aes(x = as.numeric(year), y = n), data = age_data_tidyr_low) + 
  geom_point(alpha = 1/5, 
             position = position_jitter(h = 0)) +
  geom_smooth(method = 'lm', color = 'red')

grid.arrange(p1,p2, ncol = 1)

age_data_2015_25to31 <- subset(age_data, age_data$`2015` > 25 & age_data$`2015` < 32)
age_data_2015_25to31$`Median age`

age_data_2015_36to43 <- subset(age_data, age_data$`2015` > 35 & age_data$`2015` < 44)
age_data_2015_36to43$`Median age`


```

