Project 5 - Prosper loan data EDA by Brian Chase
========================================================

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

#Introduction

The data I explore below is Loan Data from Prosper last updated 03/11/2014. 
It contains 113,937 loans with 81 variables on each loan. 

I picked this data set since I am interested in financial data sets and 
the challenge of finding how to get a good return or---in this case---finding
the highest performing loans that do not default. 

It turned into a supervised machine learning problem since it hinges on
predicting what variables are important in predicting delinquencies.

In the end, I focused on the other side of the loan market. What influences 
Borrower APR and what are the differences in the different scoring systems 
that I discovered.

I focused on variables that would describe the borrower at the time of 
applying for the loan. For example, what is the credit score or how many 
delinquencies did the borrower have---in contrast to data collected after the
loan like the number of investors. 

I neglected variable LoanOriginalAmount originally and if I would return to the 
analysis, I'd explore that variable in more detail.

```{r}
setwd("/Users/brianchase/R/final_project")
pld <- read.csv('prosperLoanData.csv')
pld$CreditGrade <- factor(pld$CreditGrade, levels = c("","AA", "A", "B", "C",
                                                       "D", "E", "HR", "NC"))
pld$ProsperRating..Alpha. <- factor(pld$ProsperRating..Alpha., 
                                    levels = c("","AA", "A", "B", "C",
                                                       "D", "E", "HR"))
str(pld)
pld_clean <- pld[-c(207,387,388),]
pld_creditgrade_subset <- head(subset(pld_clean, CreditGrade != ""), 100)
pld_prosperrating_subset <- head(subset(pld_clean, ProsperRating..Alpha. != ""), 100)
pld_subset <- rbind(pld_creditgrade_subset, pld_prosperrating_subset)

pld_clean_proccessed <- pld_clean
pld_clean_proccessed['NumberOfLoans'] <- rep(1, nrow(pld_clean_proccessed))
pld_clean_proccessed['Index'] <- 1:nrow(pld_clean_proccessed)
pld_clean_proccessed <- pld_clean_proccessed[c("Index", "NumberOfLoans", "BorrowerAPR",  "CreditScoreRangeUpper", "CreditGrade", "ProsperRating..Alpha.")]
pld_clean_proccessed$CreditScoreCat<-cut(pld_clean_proccessed$CreditScoreRangeUpper, c(0,550,600,640,680,720,760,900))
pld_clean_proccessed$ProsperRatingCat<-cut(pld_clean_proccessed$BorrowerAPR, c(0,0.1,0.16,0.2,0.25,0.3,0.35,0.6))
pld_clean_proccessed <- pld_clean_proccessed[complete.cases(pld_clean_proccessed),]
write.csv(pld_clean_proccessed, "prosperLoanDataFull.csv")


pld_subset['NumberOfLoans'] <- rep(1, nrow(pld_subset))
pld_subset['RatingCategory'] <- rep("", nrow(pld_subset))
pld_subset['Index'] <- 1:nrow(pld_subset)

pld_subset$RatingCategory[pld_subset$CreditGrade != ""] <- "CreditGrade"
pld_subset$RatingCategory[pld_subset$ProsperRating..Alpha. != ""] <- "ProsperRating"
pld_subset <- pld_subset[c("Index", "NumberOfLoans", "RatingCategory", "BorrowerAPR",  "CreditScoreRangeUpper", "CreditGrade", "ProsperRating..Alpha.")]

pld_mincreditgrade <- tapply(pld_subset$BorrowerAPR, pld_subset$CreditGrade, min)
pld_mincreditgrade

pld_subset$CreditScoreCat<-cut(pld_subset$CreditScoreRangeUpper, c(0,560,600,640,680,720,760,840))

pld_subset$ProsperRatingCat<-cut(pld_subset$BorrowerAPR, c(0,0.1,0.16,0.2,0.25,0.3,0.35,0.6))

write.csv(pld_subset, "prosperLoanData200a.csv")

pld$CreditGradeNum <- as.numeric(factor(pld$CreditGrade, levels=c("AA", "A", "B", "C",
                                                       "D", "E", "HR", "NC")))

summary(pld$CreditGrade)
table(pld$CreditGrade)
tapply(pld$BorrowerAPR, pld$ProsperRating..Alpha., summary)
tapply(pld$CreditScoreRangeUpper, pld$CreditGrade, summary)
summary(pld$ProsperRating..Alpha.)

table(pld_subset$CreditScoreCat, pld_subset$CreditGrade)
table(pld_subset$ProsperRatingCat, pld_subset$ProsperRating..Alpha.)

pld_p2 <- pld_clean_proccessed %>% 
  group_by(ProsperRating..Alpha., CreditGrade, CreditScoreCat) %>%
  summarize(n=n())

colnames(pld_p2)[4] <- "NumberOfLoans"
write.csv(pld_p2, "prosperLoanDataSmall.csv")
```

The data includes 81 variables and 113937 observations.

```{r}
pld_12variables <- pld[c("ProsperRating..Alpha.", "ProsperRating..numeric.",
                         "ProsperScore", "LoanStatus", "CreditGrade",
                         "BorrowerAPR", "Term" , "ListingCategory..numeric.",
                         "Occupation", "OpenCreditLines", 
                         "InquiriesLast6Months", "DelinquenciesLast7Years", 
                         "RevolvingCreditBalance", "DebtToIncomeRatio", "LoanOriginalAmount")]
```


# Univariate plots section

```{r}
library(ggplot2)
qplot(x = CreditGrade, data = pld )
summary(pld$CreditGrade)
```

A critical variable for a loan is the credit grade. Here we can see that a
large amount of the data is unlabled. That is because they used a different
variable after July 2009. Let's look at the 'credit grade' equivalent after
July 2009 which is ProsperRating.

```{r}
qplot(x = ProsperRating..Alpha., data = pld )
summary(pld$ProsperRating..Alpha.)
```

The histogram for prosper rating also has missing values. My hypothesis is
that it is the data points that were accounted by CreditGrade. Next I'll 
remove all the grades/ratings that have the value "" (not labeled)

```{r}
qplot(x = CreditGrade, data = subset(pld, CreditGrade != "") )
qplot(x = ProsperRating..Alpha., 
      data = subset(pld, ProsperRating..Alpha. != "") ) 
```

Here we see for both time periods the grade/rating with the highest frequency
is the 'C' rating. 

It also looks like the ratio to higher grade/rating is skewed toward higher 
quality loans, but it is hard to tell since the ProsperRating variable has
a lower ratio of AA rated loan notes. They may have also changed the
requirements for the different classifications when they changed the variable
name. I'd have to do more research to check if they can be compared.


```{r}
library(dplyr)
#library(plyr)
#detach("package:plyr", unload=TRUE)
prosperrating_total <- count(subset(pld, ProsperRating..Alpha. != ""), 
                             'ProsperRating..Alpha.')

pld %>% 
  group_by(ProsperRating..Alpha.) %>%
  summarise(no_rows = length(ProsperRating..Alpha.),
            ratio_pct = length(ProsperRating..Alpha.) / 
              prosperrating_total$n * 100)
```

Here the percentage of notes above C is 41.84

```{r}

creditgrade_total <- count(subset(pld, CreditGrade != ""), 'CreditGrade')

subset(pld, CreditGrade != "") %>% 
  group_by(CreditGrade) %>%
  summarise(no_rows = length(CreditGrade),
            ratio_pct = length(CreditGrade) / creditgrade_total$n * 100)
```

And for credit grade, the notes above C is 38.73. So my visual intuition was
correct---but the difference is not much.

```{r}
table(pld$Term)
summary(pld$Term)
```

There are three values for Term. 12 months, 36 and 60 months

```{r}
qplot(x = LoanStatus, data = pld ) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

loanstatus_total <- count(pld, 'LoanStatus')

pld %>% 
  group_by(LoanStatus) %>%
  summarise(no_rows = length(LoanStatus),
            ratio_pct = length(LoanStatus) / loanstatus_total$n * 100)

summary(pld$LoanStatus)
```

In the LoanStatus variable, we can see that most loans are current and the
majority of the ended loans are completed. But there is an interesting level
of chargedoff and defaulted loans. When you add chargedoff and Defaulted
variable counts together, it is shy of half the completed variable count. 
This seems to be a high number for unsecured loans. Does grade matter when
it comes to what is completed and what is chargedoff/defaulted?


```{r}
qplot(x = BorrowerAPR, data = pld, binwidth = 0.01)
summary(pld$BorrowerAPR)
```


```{r}
qplot(x = ListingCategory..numeric., data = pld, binwidth = 1) +
  scale_x_continuous(breaks = seq(0, 20, 1)) +
  coord_flip() 

listingcategory_total <- count(pld, 'ListingCategory..numeric.')

pld %>% 
  group_by(ListingCategory..numeric.) %>%
  summarise(no_rows = length(ListingCategory..numeric.),
            ratio_pct = length(ListingCategory..numeric.) / 
              listingcategory_total$n * 100) %>%
  arrange(desc(no_rows))
```

From variable explaination spreadsheet:
"The category of the listing that the borrower selected when posting their 
listing: 0 - Not Available, 1 - Debt Consolidation, 2 - Home Improvement,
3 - Business, 4 - Personal Loan, 5 - Student Use, 6 - Auto, 7- Other, 
8 - Baby&Adoption, 9 - Boat, 10 - Cosmetic Procedure, 11 - Engagement Ring, 
12 - Green Loans, 13 - Household Expenses, 14 - Large Purchases, 
15 - Medical/Dental, 16 - Motorcycle, 17 - RV, 18 - Taxes, 19 - Vacation, 
20 - Wedding Loans"

It looks like debt considation is by far and away the largest category. 
After that is NA and Other. Then there is Home improvement loans and business 
loans.

```{r}
qplot(x = Occupation, data = pld ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

summary(pld$Occupation)

pld %>% 
  group_by(Occupation) %>%
  summarise(no_rows = length(Occupation)) %>%
  arrange(desc(no_rows))
```

Excluding Other, the top occupations applying for loans is Professional---which 
could encompass some of other occupations. Then the next is Computer programmer
and exective. It'd be interesting to see if professions were skewed toward 
different types of loans.

```{r}
qplot(x = OpenCreditLines, data = pld, binwidth = 0.1 ) + 
  scale_x_continuous(limits = c(0, 30))

summary(pld$OpenCreditLines)
```

The open credit lines peaks around 7 credit lines. The curve almost looks like
a normal distribuition with a long tail.


```{r}
qplot(x = InquiriesLast6Months, data = pld) + 
  scale_x_continuous(limits = c(0, 25), breaks = seq(0, 25, 1)) 

summary(pld$InquiriesLast6Months)
```

For inquires in the last 6 months, most fall with 0 to 3. But someone did have
105 inquires!

```{r}
qplot(x = DelinquenciesLast7Years, data = pld )

summary(pld$DelinquenciesLast7Years)
```

```{r}
qplot(x = RevolvingCreditBalance, data = pld ) +
  scale_x_continuous(limits = c(0, 50000))

summary(pld$RevolvingCreditBalance)
```

The revolving credit balance looks like a power law distubution.  

```{r}
qplot(x = DebtToIncomeRatio, data = pld ) +
  scale_x_continuous(limits = c(0, 1))

summary(pld$DebtToIncomeRatio)
```

Most loan applicants have a debt-to-incomeRatio from 0.15 to 0.25. The peak is
around 0.25. 

```{r}
qplot(x = BorrowerAPR, data = pld, binwidth = .001) +
  scale_x_continuous(limits = c(0.05, 0.4))

qplot(x = BorrowerAPR, data = pld, binwidth = .01) +
  scale_x_continuous(limits = c(0.05, 0.4)) +
  facet_wrap(~ProsperRating..Alpha., scales="free_y")

summary(pld$BorrowerAPR)
```

BorrowerAPR median is 21% (rounded) with a mean of 21.9%. 

```{r}
qplot(x = LoanOriginalAmount, data = pld, bins = 25) +
  scale_x_log10()
```

LoanOriginalAmount almost looks like a normal distribution when transformed into
a logrithmic scale---there are spikes where there are popular values.

# Univariate Analysis

### What is the structure of the dataset?
There are 81 variables and 113937 observations.

(worst) —————-> (best)

CreditRating: HR,E,D,C,B,A,AA

ProsperRating: HR,E,D,C,B,A,AA

Other observations:
* 15% of notes are in the defaulted or chargedoff loan status
* 51% of loans are listed in the Debt Consildation category.
* The occupation listed on loans is usually "professional"
* Most loan applicants have a debt-to-incomeRatio from 0.15 to 0.25. The peak
is around 0.25.


In terms of credit score, the middle rating/grade "C" is the most frequent

### What is/are the main features(s) of interest in the dataset?

The main features are the CreditGrade/ProsperRating/ProsperScore
(which one to use?!), Loan Status, and BorrowerAPR. 

Lending club is a peer-to-peer lending marketplace where loan applicants are
trying to get the lowest APR and the investors are trying to make the highest
return---which comes down to figuring out the rate that unsecured loans will 
written off/defaulted. 

# Bivariate Plots Section

```{r}
qplot(x = ProsperScore, data = pld ) +
  facet_wrap(~ProsperRating..Alpha., scales="free_y")
summary(pld$ProsperScore)
```

The chart above is faceted based on the ProsperRating..Alpha variable.

ProsperScore also measures the metric of risk (higher = less risk). It is
interesting that the ProsperScores are varied for each ProsperRating. I was
expecting that it would be a copy of ProsperRating. Why is this different?

```{r}
qplot(x = ProsperScore, data = pld ) +
  facet_wrap(~CreditGrade, scales="free_y")
```

There are only ProsperScores during the time period after CreditGrade...

```{r} 
library(psych)
pld_12variables_numeric <- pld_12variables[sapply(pld_12variables, is.numeric)]

lowerCor(pld_12variables[sapply(pld_12variables, is.numeric)])
corPlot(pld_12variables[sapply(pld_12variables, is.numeric)])

pairs.panels(pld_12variables_numeric)
```

Going to look for the top 15 correlations for ProsperScore,
ProsperRating..numeric., and Borrowing APR and see if there are any variables
that I've been missing in a list.

```{r}
z <- cor(pld[, 16], pld[ , sapply(pld, is.numeric)], use = "pairwise.complete.obs")
z <- as.data.frame(as.table(z))
head(z[order(-abs(z$Freq)),], 15)
```

InquiriesLast6Months, CreditScoreRange, and investors variables look to be 
variables that most likely is not derived from other variables that correlate
well with ProsperScore.

```{r}
z <- cor(pld[, 82], pld[ , sapply(pld, is.numeric)], use = "pairwise.complete.obs")
z <- as.data.frame(as.table(z))
head(z[order(-abs(z$Freq)),], 20)
```

I could see that BorrowerAPR and ProsperRating..numeric. correlated highly from
the previous tables above, so I plotted just the ProsperRating correlations and 
sorted them. CreditScoreRange, Investors (I am guessing this means the size of 
the loan since most people put in $25 increments to diversify---not going to 
count this on second thought), and Inquiries (both Total and the last 6 months). 

```{r}
z <- cor(pld[, 14], pld[ , sapply(pld, is.numeric)], use = "pairwise.complete.obs")
z <- as.data.frame(as.table(z))
head(z[order(-abs(z$Freq)),], 20)
```

```{r}
qplot(x = BorrowerAPR, y = InquiriesLast6Months, data = pld)
```

It looks like there are a few outliers that extend far away from the other
points (~100 inquiries). It is interesting that there are a few vertical lines.
I wonder why that is. I want to see if I can get a closer look at the data 
excluding the outliers.

```{r}
ggplot(aes(x = BorrowerAPR, y = InquiriesLast6Months), data = pld) + 
  geom_jitter(alpha = 1/25) + 
  xlim(0.05, 0.425) +
  ylim(0, 20)
```

People have fewer inquiries to the lower end of the BorrowerAPR range 
(better credit), but it is rather uniform---with larger peaks in short bands at
the very end. Also there is a strange gap around 0.36.

```{r}
ggplot(aes(x = BorrowerAPR, y = InquiriesLast6Months), data = pld) + 
  geom_jitter(alpha = 1/5) + 
  xlim(0.30, 0.39) +
  ylim(0, 20)
```

There looks to be less points between x = 0.36 to 0.375. 

```{r}
qplot(x = LoanStatus, y = BorrowerAPR, 
      data = pld,
      geom = 'boxplot') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

The median BorrowerAPR for Completed is the lowest of the LoanStatus categories.
They are as I would expect. The PastDue(>120 days) category is interesting 
in that there is a larger jump in median borrower APR and a tighter range
favoring higher APR. Could this be because the other 'less past due' notes jump 
to Defaulted or Chargedoff through some action on the borrower and higher 
borrowerAPR borrowers tend to negliect that action?

#Bivariate Analysis

###How did the feature(s) of interest vary with other features in the dataset?
ProsperRating, BorrowerAPR, LenderYield, BorrowerRate, EstimatedLoss are highly
correlated (greater than the absolute value of 0.85).

ProsperScore, CreditScoreRangeLower/Upper are slightly less correlated 
(greater than the absolute value of 0.65).

This leads one to question which variables are derived from other variables. 
It makes sense that they use the ProsperRating to determine the BorrowerAPR to
a large extent and the Lender's yield obvious is critically tied to what the
borrower pays in interest rate---unless the Prosper the platform company
charged wildly variying rates based on a different variable.

In comparing LoanStatus to the BorrowerAPR, there is lower median APR for
completed and current loan notes when compared to the ones that are
chargedoff/defaulted/past due.


# Multivariate Plots Section

```{r}
ggplot(aes(x = BorrowerAPR, y = DelinquenciesLast7Years, color = CreditGrade),
       data = subset(pld, CreditGrade != "")) +
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  xlim(0, 0.4) +
  scale_color_brewer() +
  theme_dark()
```

Delinquencies are less for smaller BorrowerAPR as expected. However, I was 
expecting the CreditGrade to match up with the borrower APR. There is a trend 
for the better CreditGrade notes being lower BorrowerAPR, but there is a great
spread and overlap.

```{r}
ggplot(aes(x = BorrowerAPR, y = DelinquenciesLast7Years, 
           color = ProsperRating..Alpha.), 
           data = subset(pld, ProsperRating..Alpha. != "")) +
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  xlim(0, 0.4) +
  scale_color_brewer() +
  theme_dark()
```

This looks more like I would expect! The ProsperRatings are tightly coupled in 
bins organized by BorrowerAPR. The Delinquencies for the last 7 years seem to
matter most for AA and A ratings. There is some overlap of ProsperRating in the
different BorrowerAPR bins, but it is much less than for Credit Grade.The bins 
are a little less defined for the high APRs.

```{r}
ggplot(aes(x = BorrowerAPR, y = DelinquenciesLast7Years, color = ProsperScore), 
       data = subset(pld, !is.na(ProsperScore))) +
  geom_point(alpha = 0.2, size = 1, position = 'jitter') +
  xlim(0, 0.4) +
  theme_dark()
```

There is an obvious trend toward darker, lower ProsperScore for higher
BorrowerAPR. There looks to be some overlap of Scores. Making the dots more
transparent, we can saee the tight distribution of deliquencies in the last 7
years for Borrower APR. 

```{r}
ggplot(aes(x = BorrowerAPR, y = CreditScoreRangeUpper, color = CreditGrade), 
       data = subset(pld, CreditGrade != "")) +
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  xlim(0, 0.4) +
  scale_color_brewer() +
  theme_dark()
```

This is clear from this chart that the CreditGrade is dominated by 
CreditScoreRangeUpper (and potentially CreditScoreRangeLower). There are clear
horizontal colored bands correlating to CreditScore bands. The HR and E 
CreditGrades bands overlap a little. 

While the BorrowerAPR is trending how I would have guessed --- AA for lower 
BorrowerAPR and E is the highest --- there is significant overlap between 
CreditGrades and BorrowerAPR. How would they decide what BorrowerAPR to give 
for someone applying for a 17% loan?

```{r}
ggplot(aes(x = BorrowerAPR, y = CreditScoreRangeUpper, 
           color = ProsperRating..Alpha.), 
           data = subset(pld, ProsperRating..Alpha. != "")) +
  geom_point(alpha = 0.1, size = 1, position = 'jitter') +
  xlim(0, 0.4) +
  scale_color_brewer() +
  theme_dark()
```

Here we can see the same bands of ProsperRating binned by BorrowerAPR. Now we 
can see how that compares to CreditScore---where there is significant overlap.

```{r}
ggplot(aes(x = BorrowerAPR, y = CreditScoreRangeUpper, color = ProsperScore), 
       data = subset(pld, !is.na(ProsperScore))) +
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  xlim(0, 0.4) +
  theme_dark()
```

Same chart colored with ProsperScore. There is a trend of lower prosperScore
for lower BorrowerAPR and lower CreditScore. 

# Final Plots and Summary

###Plot One

```{r}
# Reordering LoanStatus levels based on median BorrowerAPR
pld2 <- pld
LoanStatusOrdered <- subset(pld, !is.na(pld$BorrowerAPR)) %>% 
  group_by(LoanStatus) %>%
  summarise(BorrowMedian = median(BorrowerAPR)) %>%
  arrange(BorrowMedian)

pld2$LoanStatus <- factor(pld2$LoanStatus, levels = LoanStatusOrdered$LoanStatus) 

qplot(x = LoanStatus, y = BorrowerAPR, 
      data = pld2,
      geom = 'boxplot') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "BorrowerAPR by Loan Status") + 
  coord_flip()
```

###Description One

Loans that are Chagedoff, Defaulted or are Past Due have higher median 
BorrowerAPR. Completed has the lowest median and the notes that are >120 days
past due have a slight uptick for median.

###Plot Two

```{r}
ggplot(aes(x = BorrowerAPR, y = CreditScoreRangeUpper, color = CreditGrade), 
       data = subset(pld, CreditGrade != "")) +
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  xlim(0, 0.4) +
  ylim(450, 900) +
  labs(y = "Credit Score Range Upper", x = "Borrower APR", 
       title = "Credit Score Range Upper by Borrower's APR faceted by 
       CreditGrade with conditional mean lines") +
  scale_color_brewer() +
  geom_smooth(aes(color = CreditGrade), se = FALSE, size = 2) +
  theme_dark() 
```

###Description Two

CreditGrade colored data points falls into well-defined buckets of CreditScore
with AA being the highest CreditScoreRangeUpper interval. In contrast, the 
variable BorrowerAPR there is a trend of lower BorrowerAPR for higher grade 
loan notes similar to Credit Score, but there are overlaps between the 
CreditGrade of the loan notes.

###Plot Three

```{r}
ggplot(aes(x = BorrowerAPR, y = CreditScoreRangeUpper, 
           color = ProsperRating..Alpha.), 
           data = subset(pld, ProsperRating..Alpha. != "")) +
  geom_point(alpha = 0.2, size = 1, position = 'jitter') +
  xlim(0, 0.4) +
  labs(y = "Credit Score Range Upper", x = "Borrower APR", 
       title = "Credit Score Range Upper by Borrower's APR faceted by 
       ProsperScore with linear smoothed conditional mean line") +
  scale_color_brewer() +
  geom_smooth(aes(color = ProsperRating..Alpha.), se = FALSE, method = "lm", size = 3) +
  theme_dark()
```

###Description Three

ProsperRating colored data points falls into well defined buckets of 
Borrower's APR with the highest rated notes AA having the lowest Borrower APR
all the way down to HR rated loans notes having the highest Borrower APR. 
In contrast, the CreditScoreRangeUpper variable has a trend of lower 
BorrowerAPR for higher rated notes similar to BorrowerAPR, but there are 
overlaps between the ProsperRating sections.

# Reflection

I ran into difficulties getting into the mindset to explore the data. My initial
thought was that I was going to find some interesting patterns in the data that 
specifically made a borrower have a good or bad APR and figure out what makes a
loan a good canidate to be completed or default. While it was a good starting
point, I started to realize that I needed to keep a more open mind, because I
was litterially feeling stressed when something profound did not jump out at me.
Looking at expected relationships is fine and can be enlightening too!

I had more success when I started allowing myself to browse the data for 
expected relationships. I felt success when I realized that CreditGrade 
depended on the CreditScore and it was so plainly visible in the charts when 
compared to a chart of the same with ProsperScore. That was satisfying.

The analysis could be further enriched by exploring the ProsperScore and how
that is created. Also I would like to create a classifier to figure out what 
variables most impact whether or not a note is a "good" loan note. I tried to 
create one in R where I binned good outcomes/most likely good outcomes and bad 
outcomes/most likely bad outcomes (half the past due notes split in each 
category based on how late they were), but my model crashed Rstudio and I did 
not follow through on debugging the reason why. The weights of that model could
open new paths to explore.